---
title: "Econ 390: Team Assignment #1 (solo try)"
author: Michelle Schultze
output: pdf_document
date: "2023-02-07"
---

Load packages & import data

```{r load-import}
library(tidyverse)
library(broom)
library(tidymodels)
library(ggplot2)
library(pls)
library(haven)

data_month <- read_csv("/Users/michelle/econ390/econ390-assignment-1/data/data_month_NL.csv")
# data_month$month <- paste('0', data_month$month, sep="")
data_month$Date <- paste(data_month$year,data_month$month,"01", sep="-")
data_month$Date <- as.Date(data_month$Date, format = "%Y-%m-%d")


regGDP <- read_dta("data/Region by year GDP panel.dta")
regGDP[regGDP$year == 2021, 3] <- NA 

```

Each team should mainly focus on their regions, but comparing with the country level (sum/average) is welcomed. 

Assigned regions:

UA_21	Zakarpatska	1
UA_26	Ivano-Frankivska	1
UA_46	Lvivska	1
UA_61	Ternopilska	1
UA_77	Chernivetska	1


```{r map}
#library(maps)
library(rgdal)
library(maptools)
#library(sf)
#library(terra)

download.file("https://geodata.ucdavis.edu/gadm/gadm4.1/shp/gadm41_UKR_shp.zip", destfile="data/ukraine_map.zip")
system("unzip data/ukraine_map.zip")

map_1 <- readOGR( 
  dsn= paste0(getwd()) , 
  layer="gadm41_UKR_1",
  verbose=FALSE, stringsAsFactors = FALSE)

plot(map_1)

str(map_1, max.level = 2)
str(map_1@data)
map_1@data$NAME_1

map_1@data  <- map_1@data %>%
  mutate(color = ifelse(NAME_1 %in% c("Zakarpattia","Ivano-Frankivs'k","L'viv","Ternopil'","Chernivtsi"), "red", "white"))

plot(map_1, col = map_1@data$color)


```

Merge these datasets together using year and region. Use left_join to merge different datasets. As for data_month_NL.csv, which contains the monthly information, you can either generate annual variables from the monthly information or try to pivot the datasets to create 12 new variables for each variable you think useful (e.g. NLI_01 for January, NLI_02 for February, NLI_03 for March, etc.).

```{r prep-data}

## code from class

data_tidy_YR_test <- data_month |> group_by(year, reg, name) |> 
  mutate(NLI = nl_sum_4/(area_sq_km - 0.141*nl_nodata_4),
         NLI2 = nl_mean_4/nl_std_4)

data_tidy_YR_test <- filter(data_tidy_YR_test, year >= 2012, !(reg %in% c("UA_09_O", "UA_14_O")) ) 

# remove LNR/DNR {careful with character string "UA_09_0" vs. "UA_14_O"} 
# original data_month starts in 2010 (with many missing variables)
data_tidy_YR_test <- filter(data_tidy_YR_test, year >= 2012, !(reg %in% c("UA_09_O", "UA_14_O")) ) 

# arrange order of columns for easier viewing
data_tidy_YR_test <- data_tidy_YR_test |> 
  select(c(year, month, reg, name, area_sq_km, NLI, NLI2, tw_count), everything())


# Coding Question:
# How do we compute logs of a select range of columns?
test <- data_tidy_YR_test
test$logNLI <- log(test$NLI)
# What happens when we take log of negative values or zero? Are the outputs the same?
# Check for -Inf and NaN
# How do we identify this?
# check is.na() function and is.infinite() function
# What do we do about that once identified?
# sum(is.na(test$logNLI))
# sum(is.infinite(test$logNLI))
# which(is.na(test$logNLI))
# which(is.infinite(test$logNLI))

data_tidy_YR_test <- data_tidy_YR_test |> 
  mutate(across(c(NLI:tw_n_settlements), ~log(.x), .names = '{.col}'))

data_tidy_YR_test[data_tidy_YR_test == -Inf] <- 0

# Estimation Portion 2012-2020 (2021 we have monthly levels)
data_tidy_YR <- filter(data_tidy_YR_test, year < 2021) |>
  group_by(year, reg, name) |>
  summarise(across(c(NLI:tw_n_settlements), ~mean(.x, na.rm=TRUE), .names = '{.col}')) |>
  mutate(across(c(NLI:tw_n_settlements), ~round(.x, digits=3), .names = '{.col}'))

# Add yearly levels of GDP
data_tidy_YR <- left_join(data_tidy_YR, regGDP, by=c("year", "reg"))


data_tidy_YR$GDP <- log(data_tidy_YR$GDP)
colnames(data_tidy_YR)[25] <- "lng"


```

Use pivot_wider() to pivot the datasets.

```{r}
# ??
```

Use these datasets to create some line charts (geom_line)

```{r line-charts}
theme_set(theme_bw())

# NL yearly data
# Coding question: 
# How do plot the evolution over time of a variable across all region?
# Check use of facet_wrap
ggplot(filter(data_tidy_YR, year <=2021), aes(x=year, y=nl_mean_4))+
  geom_smooth(method = "lm") + facet_wrap(~name, scales = 'free_y') + 
  ylab("Average Luminosity") +
  geom_rect(data = subset(data_tidy_YR, name %in% c("Zakarpatska","Ivano-Frankivska","Lvivska","Ternopilska","Chernivetska")), 
                          fill = NA, colour = "red", linewidth = 2, xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggplot(filter(data_tidy_YR, year <=2021, name %in% c("Zakarpatska","Ivano-Frankivska","Lvivska","Ternopilska","Chernivetska")), aes(x=year, y=nl_mean_4))+
  geom_smooth(method = "lm", aes(color=name), show.legend = FALSE) + facet_wrap(~name, scales = 'free_y') + 
  ylab("Average Luminosity") + 
  geom_point(aes(color=name), show.legend = FALSE) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Mean Night Light Luminosity", subtitle = "Western Ukraine")

## what would the title be? what is nl_mean_4? 
```



```{r line-charts-2}
data_tidy_YR <- data_tidy_YR %>%
  mutate(of_interest = ifelse(name %in% c("Zakarpatska","Ivano-Frankivska","Lvivska","Ternopilska","Chernivetska"), TRUE, FALSE))

ggplot(filter(data_tidy_YR, year <=2020), aes(x=log(nl_mean_4), y=lng))+
  geom_point(aes(color = of_interest)) + geom_smooth(method="lm")

# lower ratio of (log) luminosity to (log) GDP in Western Ukraine

ggplot(data = filter(data_tidy_YR, year <=2020), aes(x=log(nl_mean_4), y=lng))+
  geom_point(aes(color=of_interest)) + geom_smooth(method="lm", color = "black") +
  geom_smooth(data = filter(data_tidy_YR, year <=2020, of_interest == TRUE), aes(x=log(nl_mean_4), y=lng), method="lm", color = "blue")


# Western regions relative to each other: 

ggplot(filter(data_tidy_YR, year <=2020, name %in% c("Zakarpatska","Ivano-Frankivska","Lvivska","Ternopilska","Chernivetska")), aes(x=log(nl_mean_4), y=lng))+
  geom_point() + geom_smooth(method="lm") + 
  geom_text(aes(label = name, color=name), check_overlap = TRUE)+
  theme(legend.position="none")
```


```{r line-graphs-3}
ggplot(data = filter(data_tidy_YR, year <=2020), aes(x=log(nl_mean_4), y=lng))+
  geom_point(aes(color=of_interest)) + geom_smooth(method="lm", color = "black", se=FALSE) +
  geom_smooth(data = filter(data_tidy_YR, year <=2020, of_interest == TRUE), aes(x=log(nl_mean_4), y=lng), method="lm", color = "blue", se=FALSE)+
  facet_wrap(~year, scales = 'free')
```




bar charts (geom_bar)

```{r}






```


pie charts (geom_bar + coord_polar())

```{r}

```


boxplots (geom_boxplot)

```{r}

```


Use these datasets to fit fixed effect line models. Use plm function to run the fixed effect regression.

```{r}

library(plm)
library(AER)
library(stargazer)


########## GDP and Nightlights Panel Models #####
# filtering oblasts with available GDP over entire period
panel_data <- filter(data_tidy_YR, !(reg %in% c("UA_40", "UA_43")))


### Unit FE Model
NL_fe_mod <- plm(lng ~ NLI,
                 data = panel_data,
                 index = c("name", "year"),
                 model = "within")
summary(NL_fe_mod)
coeftest(NL_fe_mod, vcov = vcovHC, type = "HC1")


# Unit FE == OLS with dummies for Units
NL_lm_mod <- lm(lng ~ NLI + name - 1, data = panel_data)
summary(NL_lm_mod)


### Time and Unit FE
NL_tefe_mod <- plm(lng ~ NLI,
                   data = panel_data,
                   index = c("name", "year"),
                   model = "within",
                   effect = "twoways")
summary(NL_tefe_mod)
coeftest(NL_tefe_mod, vcov = vcovHC, type = "HC1")

# Time and Unit FE == OLS with dummies for Units and Years
NL_TU_lm_mod <- lm(lng ~ NLI + name + as.factor(year) - 1, data = panel_data)
summary(NL_TU_lm_mod)


# Arranging results for presentation
model_se <- list(sqrt(diag(vcovHC(NL_fe_mod, type = "HC1"))),
                 sqrt(diag(vcovHC(NL_lm_mod, type = "HC1"))),
                 sqrt(diag(vcovHC(NL_tefe_mod, type = "HC1"))),
                 sqrt(diag(vcovHC(NL_TU_lm_mod, type = "HC1"))))

# output latex table (type = "html" for HTML/CSS code)
stargazer(NL_fe_mod, NL_lm_mod, NL_tefe_mod, NL_TU_lm_mod,
          digits = 3,
          header = FALSE,
          type = "html",
          se = model_se,
          title = "Linear Panel Regression Models of GDP and Nightlights",
          model.numbers = FALSE,
          column.labels = c("(1)", "(2)", "(3)", "(4)"),
          out = "table.htm")

## how to make table come out fully formatted?


```


Try to go through the model selection methods (Best Subset Selection - regsubsets, Ridge – glmnet alpha = 1, LASSO – glmnet alpha = 0) to find the best prediction model for GDP. Start with as many variables as you can get.

```{r linear}

## class code



## couldn't figure out how to make regfit.full work, error: "Exhaustive search will be S L O W, must specify really.big=T"

###
library(ISLR2)
dim(data_tidy_YR)
sum(is.na(data_tidy_YR$lng))
sum(is.na(data_tidy_YR$nl_mean_4))
###
data_tidy_YR_2 <- na.omit(data_tidy_YR)
dim(data_tidy_YR_2)
sum(is.na(data_tidy_YR_2))
###
library(leaps)
regfit.full <- regsubsets(nl_mean_4 ~ ., data_tidy_YR_2, nvmax = 5)
summary(regfit.full)




```

```{r}
## Ridge Regression and the Lasso

###
x <- model.matrix(nl_mean_4 ~ ., data_tidy_YR_2)[, -1]
y <- data_tidy_YR_2$nl_mean_4

### Ridge Regression

###
library(glmnet)
grid <- 10^seq(10, -2, length = 100)
ridge.mod <- glmnet(x, y, alpha = 0, lambda = grid)
###
dim(coef(ridge.mod))
###
ridge.mod$lambda[50]
coef(ridge.mod)[, 50]
sqrt(sum(coef(ridge.mod)[-1, 50]^2))
###
ridge.mod$lambda[60]
coef(ridge.mod)[, 60]
sqrt(sum(coef(ridge.mod)[-1, 60]^2))
###
summary(predict(ridge.mod, s = 50, type = "coefficients")[1:20, ])
###
set.seed(1)
train <- sample(1:nrow(x), nrow(x) / 2)
test <- (-train)
y.test <- y[test]
###
ridge.mod <- glmnet(x[train, ], y[train], alpha = 0,
                    lambda = grid, thresh = 1e-12)
ridge.pred <- predict(ridge.mod, s = 4, newx = x[test, ])
mean((ridge.pred - y.test)^2)
###
mean((mean(y[train]) - y.test)^2)
###
ridge.pred <- predict(ridge.mod, s = 1e10, newx = x[test, ])
mean((ridge.pred - y.test)^2)
###
ridge.pred <- predict(ridge.mod, s = 0, newx = x[test, ],
                      exact = T, x = x[train, ], y = y[train])
mean((ridge.pred - y.test)^2)
lm(y ~ x, subset = train)
predict(ridge.mod, s = 0, exact = T, type = "coefficients",
        x = x[train, ], y = y[train])[1:20, ]
###
set.seed(1)
cv.out <- cv.glmnet(x[train, ], y[train], alpha = 0)
plot(cv.out)
bestlam <- cv.out$lambda.min
bestlam
###
ridge.pred <- predict(ridge.mod, s = bestlam,
                      newx = x[test, ])
mean((ridge.pred - y.test)^2)
###
out <- glmnet(x, y, alpha = 0)
predict(out, type = "coefficients", s = bestlam)[1:20, ]


## what are these numbers describing, what would be useful to look at here? 
## should we highlight and discuss the coefficients for our regions of interest? 




### The Lasso

###
lasso.mod <- glmnet(x[train, ], y[train], alpha = 1,
                    lambda = grid)
plot(lasso.mod)
###
set.seed(1)
cv.out <- cv.glmnet(x[train, ], y[train], alpha = 1)
plot(cv.out)
bestlam <- cv.out$lambda.min
lasso.pred <- predict(lasso.mod, s = bestlam,
                      newx = x[test, ])
mean((lasso.pred - y.test)^2)
###
out <- glmnet(x, y, alpha = 1, lambda = grid)
lasso.coef <- predict(out, type = "coefficients",
                      s = bestlam)[1:20, ]
lasso.coef
lasso.coef[lasso.coef != 0]



## same questions here

```


